Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-19 eventPointTwoLoop.tex
added citations and clarified exponentiations

\documentclass[11pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scalebox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scalebox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

@@

\newcommand{\simplebox}{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-1,1){2};
\vertex (a2) at (1,-1){4};
\vertex (a3) at (1,1){3};
\vertex (a4) at (-1,-1){1};
\vertex (mid3) at (.5,.5);
\vertex (mid4) at (.5,-.5);
\vertex (mid5) at (-.5,.5);
\vertex (mid6) at (-.5,-.5);
\diagram{
(a4) --[ultra thick,](mid6),
(a3) --[ultra thick,](mid3),
(a2) --[ultra thick,](mid4),
(a1) --[ultra thick,](mid5),
(mid3) --[ultra thick,](mid4),
(mid5) --[ultra thick,](mid6),
(mid5) --[ultra thick,](mid3),
(mid4) --[ultra thick,](mid6),
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\scaleIntAvectorODD}[6]{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a5) at (1.3,.7) {#6};
\vertex (a6) at (0,.7) {#5};
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey4](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[photon, ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[photon, ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

In what follows, we
\fi
In this section we provide a catalog of the even-point effective field theories studied in this paper. We will define the Lagrangian, $\mathcal{L}$, to be the integrand of the action, $S$, when working in spacetime coordinate frame:
\begin{equation}
S= \int d^D x \,\mathcal{L}
\end{equation}
Some theories in this work, like Volkov-Akulov and supersymmetric extensions of Born-Infeld, have natural formulations in terms of veirbeins that depend on fermionic degrees of freedom. For such theories, it is useful to distinguish between the coordinate independent actions, and the coordinate-dependent Lagrangians. 
\paragraph{Nonlinear Sigma Model}
The simplest effective field theory that we will consider in this work is the {nonlinear sigma model} (NLSM). The NLSM Lagrangian can be expressed in terms of the chiral current, $j_\mu = U^\dagger \partial_\mu U$, as follows,
\begin{equation}\label{eq:NLSMLag}
\mathcal{L}^{\text{NLSM}}= \frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] = \frac{1}{2}\text{tr}\left[\frac{(\partial_\mu\pi )( \partial^\mu \pi) }{(1-f_\pi^{-2}\pi^2)^2}\right]
\end{equation}
where the trace is over the color indices of the gauge group, $\pi \equiv \pi^a T^a$, and on the right hand side we have applied the Cayley parameterization to the $SU(N_c)$ group elements, $U= \frac{1+\pi/f_\pi}{1-\pi/f_\pi}$, where $f_\pi$ is the dimensionful pion decay width. This theory is the unique two-derivative EFT that is invariant under constant shifts of the pion field in color space:
\begin{equation}\label{eq:shiftSym}
\pi^a \rightarrow \pi^a + c^a
\end{equation}

We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf,Kampf:2013vha}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)]\text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Considering the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $SU(2)$ NLSM have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. In doing so, we show that the IR divergences of the theory exponentiates in the large $N_c$ limit. 
\paragraph{Born-Infeld Photons and Dirac Scalars}
The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory
\begin{equation}
 \label{biLag}
  \alpha'^{\,2}\mathcal{L}^{\text{BI}} = 1-\sqrt{\text{det}(\eta_{\mu\nu}+\alpha' F_{\mu\nu})} \,.
\end{equation}
A close cousin of Born-Infeld theory is the dimensionally reduction to Dirac scalars, which describes a $D$-dimensional spacetime propagating in a $(D+1)$-dimensional background:
\begin{equation}
 \label{DBI}
  \alpha'^{\,2}\mathcal{L}^{\text{DBI}} =1- \sqrt{\text{det}(\eta_{\mu\nu}+\alpha'^{\,2} \partial_\mu \varphi \partial_\nu \varphi)}
\end{equation}
where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by considering the most general polynomial  of $X=(\partial \varphi)^2$, $P(X)$,  constrained to be invariant under the field redefinition \cite{Cheung:2016drk}:
\begin{equation}\label{DBIShift}
\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)
\end{equation}

When the transformation law leaves the polynomial $P(X)$ theory invariant, there are an additional $D+1$ conserved charges, and the Poincar\'{e} symmetry of $P(X)$ EFT is promoted from $D$ to $(D+1)$ dimensions. One can verify that applying \eqn{DBIShift} to \eqn{DBI} shifts the Lagrangian by a total derivative~\cite{deRham:2010eu}. Similar to the soft theorem constraints on the building blocks of NLSM and $\chi \text{PT}$, the DBI action can be bootstrapped by requiring the amplitudes vanish at $\mathcal{O}(p^2)$ when taking external momentum, $p$, soft~\cite{Cheung:2016drk}.

@@

We begin with the color-dressed NLSM tree amplitudes that will serve as the kinematic building blocks to appear in our unitarity construction. For our purposes, a convenient color basis is the so-called half-ladder basis of Del Duca, Dixon and Maltoni (DDM) \cite{DixonMaltoni},
\begin{equation}
\mathcal{A}^{\text{NLSM}}_n = \sum_{\sigma \in S^{n-2}}\cHL_{(1\sigma n)} A^{\text{NLSM}}_{(1\sigma n)}\,.
\end{equation}
The half-ladder color factors are defined in terms of the antisymmetric structure constants as, 
\begin{equation}
\cHL_{(1\sigma n)} \equiv f^{1\sigma_2 \beta_2}f^{\beta_2 \sigma_3 \beta_3} \cdots f^{\beta_{n-1}\sigma_{n-1}n}.
\end{equation} 
In the above expression, each half ladder dresses a color-ordered amplitude $A^{\text{NLSM}}_{(1\sigma n)}$. At four-point, it is well known that the ordered amplitude for NLSM is simply, 
\begin{equation}
A^{\text{NLSM}}_{(1234)} = f_\pi^{-2} \,s_{13}\,.
\end{equation}
In which case, the full four-point amplitude can be expressed completely in terms of the $s$-channel and $u$-channel color factors, 
\begin{equation}\label{eq:NLSMamp} 
\mathcal{A}^{\text{NLSM}}_{(1|23|4)} = f_\pi^{-2}\left(\cHL_{(1234)} s_{13}+\cHL_{(1324)}s_{12}\right)\,.
\end{equation}
The $t$-channel color factor satisfies a Jacobi identity with the $s$- and $u$-channel diagrams, $c_t \equiv\cHL_{(2341)} =\cHL_{(1234)}-\cHL_{(1324)}$, and thus can be chosen so that it does not contribute to the full color-dressed amplitude. In this work, we use the convention $s=(k_1+k_2)^2$, $t=(k_2+k_3)^2$ and $u=-s-t$.  The color-dressed amplitude is entirely bose-symmetric, but we use the subscript $(1|23|4)$ to emphasize a functional choice of the $1|\sigma|4$ half-ladder basis.

@@

@@

\paragraph{\textbf{Integration}} While there is plenty of physics to be extracted from the combination of \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we will focus on the behavior of the leading divergence for $D=2-2\epsilon$ in the planar limit where $N_c \rightarrow \infty$. In this limit, the cross-box is subleading in the $N_c$ expansion, and can be neglected. In a trace basis, the double-box can be expanded as follows:

@@

C^{\text{2box}}_{(12|34)} \xrightarrow{N_c\rightarrow\infty } N_c^2\, [C(1234)+C(1432)] +\mathcal{O}(N_c)

@@

where we have defined the color trace factor, $C(abcd)\equiv  \text{tr}[T^aT^b T^c T^d]$. Furthermore, the leading divergences for the loop integrals in $D=2-2\epsilon$ dimensions are the following:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we can extract the full leading logarithmic divergence from the color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

\mathcal{A}^{\text{NLSM}}_{\text{2-loop}} = \left[C(1234)A^{2-2\epsilon}_{(1234)}+\text{cyc}(2,3,4)\right] + \mathcal{O}(N_c)

@@

which takes on the following value upon evaluating the integrals in $D=2-2\epsilon$,
A^{2-2\epsilon}_{(1234)} = -\frac{1}{2}\left[\frac{f_{\pi}^{-2}N_c }{2\pi \epsilon}\right]^2 \,f_\pi^{-2}s_{13} 

@@

We write it in this suggestive form to emphasize that the logarithmic divergence present at one-loop in \eqn{eq:1loopPionD2} appears to exponentiate! More concretely, through explicit calculation we have demonstrated that through two-loop, the leading divergence of the color ordered amplitudes in the planar limit goes as follows:
A^{2-2\epsilon}\bigg|_{\text{div.}} = A^{\text{tree}}\left( 1 + \frac{A^{\text{1-loop}}}{ A^{\text{tree}}} + \frac{1}{2}\left[\frac{A^{\text{1-loop}}}{ A^{\text{tree}}} \right]^2 + \cdots \right)
It would be interesting to investigate whether this exponential structure also can be applied to the subleading logarithms in the planar limit, similar to what was found in \cite{Anastasiou:2003kj,Bern:2005iz} in $D=4-2\epsilon$ for planar $\mathcal{N}=4$ sYM. The iterated structure of planar $\mathcal{N}=4$ has been linked to the integrability of the theory \cite{Sterman:2002qn,Minahan:2002ve,Bena:2003wd,Beisert:2003jj,Beisert:2003jb,Beisert:2003yb,Dolan:2003uh,Arutyunov:2003rg,Ryzhov:2004nz,Frolov:2005iq}. We find compelling the possibility that this iterated structure of NLSM amplitudes in $D=2-2\epsilon$ could indicate similar integrable behavior, as studied in \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}. We leave this as a direction of future study.

@@

\mathcal{L}^{\text{BI}+\text{CT}} = \mathcal{L}^{\text{BI}} + \frac{(\alpha')^4}{(4\pi)^2}\frac{1}{60}\left(\mathcal{O}^{4+}+ a_{({\text{ev.}})}\mathcal{O}^{\text{ev.}}\right)

@@

\mathcal{M}^{\text{BI+CT}}_{\text{1-loop}}\big|_{\alpha'^6} = \frac{1}{2}\frac{\alpha'^6}{60 (4\pi)^2}\scaleIntAvectorODD{}{}{}{}{$\mathcal{T}^{4+}$}{$t_8F^4$}+\frac{a_{\text{ev.}}}{2}\scaleIntAvectorODD{}{}{}{}{$\mathcal{T}^{\text{ev.}}$}{$t_8F^4$}+\text{perms}

@@

Given the particular numerical value computed in the previous section, we find evanescent Wilson coefficient must take the following $D$-dependent value:

@@

\mathcal{L}^{\text{BI}+\text{CT}} = \mathcal{L}^{\text{BI}} + \frac{\alpha'^4}{(4\pi)^2}\frac{1}{60}\left[\mathcal{O}^{4+}+ \frac{64}{(D-4)}\mathcal{O}^{\text{ev.}}\right]+\mathcal{O}(\alpha'^6)

>>>>

\documentclass[11pt,letter]{article}
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{shuffle}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\def\RA{\rlap{\scaleBox{1.6}{$\leftrightarrow$}}}
\def\DA{\smash{\bclap{\scaleBox{1.6}{$\downarrow$}}}}
\def\mystrut{\rule[-2ex]{0ex}{6ex}}

@@

\newcommand{\simpleBox}{ {
\begin{tikzpicture}[baseline=(current  bounding  box.center)]
\begin{feynman}
\vertex (a1) at (-1,1){2};
\vertex (a2) at (1,-1){4};
\vertex (a3) at (1,1){3};
\vertex (a4) at (-1,-1){1};
\vertex (mid3) at (.5,.5);
\vertex (mid4) at (.5,-.5);
\vertex (mid5) at (-.5,.5);
\vertex (mid6) at (-.5,-.5);
\diagram{
(a4) --[ultra thick,](mid6),
(a3) --[ultra thick,](mid3),
(a2) --[ultra thick,](mid4),
(a1) --[ultra thick,](mid5),
(mid3) --[ultra thick,](mid4),
(mid5) --[ultra thick,](mid6),
(mid5) --[ultra thick,](mid3),
(mid4) --[ultra thick,](mid6),
};
\end{feynman}
\end{tikzpicture}
}
}

@@

\newcommand{\scaleIntAvectorODD}[6]{ {
\begin{tikzpicture}[baseline=0]
\begin{feynman}
\vertex (a5) at (1.3,.7) {#6};
\vertex (a6) at (0,.7) {#5};
\vertex (a1) at (-.8, -0.6) {#1};
\vertex (a2) at (-.8, 0.6) {#2};
\vertex [dot,scale=3](mid1) at (0,0){};
\vertex [dot,scale=2.5,hgrey4](mid2) at (0,0){};
\vertex [dot,scale=3](mid3) at (1,0){};
\vertex [dot,scale=2.5,hgrey0](mid4) at (1,0){};
\vertex (a3) at (1.8, .6) {#3};
\vertex (a4) at (1.8, -.6) {#4};
\diagram{
(mid1) --[photon, ultra thick,](a1),
(mid1) --[photon, ultra thick,](a2),
(mid1) --[photon, ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[photon, ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[photon, ultra thick](a4),
(mid3) --[photon, ultra thick,](a3)
};
\end{feynman}
\end{tikzpicture}
}
}

@@

In what follows, we
\fi
In this section we provide a catalog of the even-point effective field theories studied in this paper. We will define the Lagrangian, $\mathcal{L}$, to be the integrand of the action, $S$, when working in spacetime coordinate frame:
\begin{equation}
S= \int d^D x \,\mathcal{L}
\end{equation}
Some theories in this work, like Volkov-Akulov and supersymmetric extensions of Born-Infeld, have natural formulations in terms of veirbeins that depend on fermionic degrees of freedom. For such theories, it is useful to distinguish between the coordinate independent actions, and the coordinate-dependent Lagrangians. 
\paragraph{Nonlinear Sigma Model}
The simplest effective field theory that we will consider in this work is the {nonlinear sigma model} (NLSM). The chiral NLSM Lagrangian with diagonal symmetry group $G=SU(N_c)$ can be expressed in terms of the chiral current, $j_\mu = U^\dagger \partial_\mu U$, as follows,
\begin{equation}\label{eq:NLSMLag}
\mathcal{L}^{\text{NLSM}}= \frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] = \frac{1}{2}\text{tr}\left[\frac{(\partial_\mu\pi )( \partial^\mu \pi) }{(1-f_\pi^{-2}\pi^2)^2}\right]
\end{equation}
where the trace is over the color indices of the gauge group, $\pi \equiv \pi^a T^a$. On the right hand side we have applied the Cayley parameterization to the $SU(N_c)$ group elements, $U= \frac{1+\pi/f_\pi}{1-\pi/f_\pi}$, where $f_\pi$ is the dimensionful pion decay width. By construction, this model has a left-right global symmetry, $G\times G$, that is nonlinearly realized, and a locally preserved diagonal subgroup, $G$. This theory is the unique two-derivative EFT that is invariant under constant shifts of the pion field in color space:
\begin{equation}\label{eq:shiftSym}
\pi^a \rightarrow \pi^a + c^a
\end{equation}
We could also consider higher derivative generalizations of the NLSM that are invariant under the shift symmetry of \eqn{eq:shiftSym}.  Such  EFTs correspond to the so-called chiral limit of {chiral perturbation theory}~\cite{Weinberg:1978kz,Gasser:1983yg,Kampf:2006yf,Kampf:2013vha}, describing massless pion scattering below the chiral symmetry breaking scale of QCD:
\begin{equation}\label{eq:XPTLag}
\mathcal{L}^{\chi\text{PT}}=\frac{1}{2}\text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)] + \frac{\beta_1}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger (\partial ^\mu U)]^2 + \frac{\beta_2}{f_\pi^2} \text{tr}[(\partial_\mu U)^\dagger( \partial_\nu U)]\text{tr}[(\partial^\mu U)^\dagger( \partial^\nu U)] +\cdots
\end{equation}
By construction, each of the higher derivative operators appearing above are invariant under the shift symmetry of \eqn{eq:shiftSym}. These higher derivative generalizations of NLSM have been studied both in the context of the $S$-matrix bootstrap \cite{Manohar:2008tc,Bellazzini:2016xrt,Guerrieri:2020bto}, and soft theorem bootstraps of refs. \cite{Cheung:2014dqa,Cheung:2015ota,Cheung:2016drk,Cheung:2018oki,Low:2019ynd}. Given the relevance of \eqn{eq:XPTLag} in Standard Model pion scattering, the amplitudes for $G= SU(2)$ isospin pions have been computed through two-loop order \cite{Bijnens:1995yn,Bijnens:1997vq,Girlanda:1997ed}. In this work, we add to the literature by computing the two-loop amplitudes of \eqn{eq:NLSMLag} for arbitrary $SU(N_c)$ color structure. We then use these amplitudes to compute $\mathcal{N}=4$ DBIVA amplitudes via the double copy in \sect{sec:DBIvDC}.

In addition to the chiral NLSM described above, our methods apply equally well to a symmetric coset model describing pion dynamics on a coset space $G/H$, with a nonlinearly realized global symmetry $G$ broken down to a linearly realized isometry group $H$. Recent literature that study the universal soft behavior of these models can be found in refs.~\cite{Low:2014nga,Low:2017mlh,Liu:2018vel,Low:2018acv}. One such model that we will study in the text is the $\mathbb{CP}^N$ model where the global $SU(N+1)$ is spontaneously broken down to a local $U(N)$ symmetry on the coset. The Lagrangian for the theory we provide below:
\begin{equation}\label{eq:CPNLag}
\mathcal{L}^{\text{NLSM}}_{\mathbb{CP}^N} = \frac{1}{2}P(z,\bar{z})^{ij}(\partial_\mu \bar{z}_i)(\partial^\mu z_j)=\frac{1}{2}\frac{(f_\pi^2+\bar{z} z)\delta^{ij}-z^i\bar{z}^j}{(f_\pi^{2}+\bar{z} z)^2}(\partial_\mu \bar{z}_i)(\partial^\mu z_j)
\end{equation}
where $z=(z_1,z_2,...,z_N)$ is a complex vector in the fundamental representation of $U(N)$, and $P(z,\bar{z})$ is the Fubini-Study metric on $\mathbb{CP}^N$. The full $SU(N+1)$ symmetry of the theory is realized due to the imbedding of $\mathbb{CP}^N$ in a complex space one dimension higher, $\mathbb{C}^{N+1}$. In \sect{sec:NLSMU}, we will demonstrate through explicit calculation that the leading IR divergences of the $\mathbb{CP}^1$ model exponentiations in $D=2$, the critical dimension of the theory. 
\paragraph{Born-Infeld Photons and Dirac Scalars}
The second even-point EFT we will study is Born-Infeld theory \cite{Born:1934gh}. This theory describes the dynamics of open string endpoints with $U(1)$ charge attached to a $D$-dimensional spacetime with Dirichlet boundary conditions. Given the stringy dynamics orthogonal to the spacetime, the electric field generated by the charged endpoints has a maximum field strength of order the inverse string tension, $1/\alpha'$. The effective Lagrangian for this theory
\begin{equation}
 \label{biLag}
  \alpha'^{\,2}\mathcal{L}^{\text{BI}} = 1-\sqrt{\text{det}(\eta_{\mu\nu}+\alpha' F_{\mu\nu})} \,.
\end{equation}
A close cousin of Born-Infeld theory is the dimensionally reduction to Dirac scalars, which describes a $D$-dimensional spacetime propagating in a $(D+1)$-dimensional background:
\begin{equation}
 \label{DBI}
  \alpha'^{\,2}\mathcal{L}^{\text{DBI}} =1- \sqrt{\text{det}(\eta_{\mu\nu}+\alpha'^{\,2} \partial_\mu \varphi \partial_\nu \varphi)}
\end{equation}
where $\varphi$ is the spacetime coordinate in the orthogonal space. This theory is uniquely determined by considering the most general polynomial  of $X=(\partial \varphi)^2$, $P(X)$,  constrained to be invariant under the field redefinition \cite{Cheung:2016drk}:
\begin{equation}\label{DBIShift}
\varphi \rightarrow \varphi + c + b^{\mu} (x_\mu + \varphi \partial_\mu \varphi)
\end{equation}
When the transformation law leaves the polynomial $P(X)$ theory invariant, there is an additional set of $D+1$ conserved charges, and the Poincar\'{e} symmetry of $P(X)$ EFT is promoted from $D$ to $(D+1)$ dimensions. One can verify that applying \eqn{DBIShift} to \eqn{DBI} shifts the Lagrangian by a total derivative~\cite{deRham:2010eu}. Similar to the soft theorem constraints on the building blocks of NLSM and $\chi \text{PT}$, the DBI action can be bootstrapped by requiring the amplitudes vanish at $\mathcal{O}(p^2)$ when taking external momentum, $p$, soft~\cite{Cheung:2016drk}.

@@

We begin with the color-dressed NLSM tree amplitudes that will serve as the kinematic building blocks to appear in our unitarity construction. For our purposes, a convenient color basis is the so-called half-ladder basis of Del Duca, Dixon and Maltoni (DDM) \cite{DixonMaltoni},
\begin{equation}
\mathcal{A}^{\text{NLSM}}_n = \sum_{\sigma \in S^{n-2}}\cHL_{(1\sigma n)} A^{\text{NLSM}}_{(1\sigma n)}\,.
\end{equation}
The half-ladder color factors are defined in terms of the antisymmetric structure constants as, 
\begin{equation}
\cHL_{(1\sigma n)} \equiv f^{1\sigma_2 \beta_2}f^{\beta_2 \sigma_3 \beta_3} \cdots f^{\beta_{n-1}\sigma_{n-1}n}.
\end{equation} 
In the above expression, each half ladder dresses a color-ordered amplitude $A^{\text{NLSM}}_{(1\sigma n)}$. At four-point, it is well known that the ordered amplitude for NLSM is simply, 
\begin{equation}
A^{\text{NLSM}}_{(1234)} = f_\pi^{-2} \,s_{13}\,.
\end{equation}
In which case, the full four-point amplitude can be expressed completely in terms of the $s$-channel and $u$-channel color factors, 
\begin{equation}\label{eq:NLSMamp} 
\mathcal{A}^{\text{NLSM}}_{(1|23|4)} = f_\pi^{-2}\left(\cHL_{(1234)} s_{13}+\cHL_{(1324)}s_{12}\right)\,.
\end{equation}
The $t$-channel color factor satisfies a Jacobi identity with the $s$- and $u$-channel diagrams, $c_t \equiv\cHL_{(2341)} =\cHL_{(1234)}-\cHL_{(1324)}$, and thus can be chosen so that it does not contribute to the full color-dressed amplitude. In this work, we use the convention $s=(k_1+k_2)^2$, $t=(k_2+k_3)^2$ and $u=-s-t$.  The color-dressed amplitude is entirely bose-symmetric, but we use the subscript $(1|23|4)$ to emphasize a functional choice of the $(1|\sigma|4)$ half-ladder basis.

@@

\begin{equation}\label{eq:NLSM1loop}

@@

\paragraph{\textbf{Exponentiation}} While there is plenty of physics to be extracted from the combination of \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we will focus on the behavior of the leading divergence for $D=2-2\epsilon$ for the symmetric coset model describing pions on $\mathbb{CP}^1\cong SU(2) /U(1)$. The tree amplitudes for this theory can be obtained from \eqn{eq:NLSMamp} by plugging in multi-trace color factors of the form,

@@

\cHL_{(ijkl)} = \delta_{ik}\delta_{jl} -\delta_{il}\delta_{jk} \,.

@@

The delta function indices select out (anti)-holomorphic pion fields that live on the $\mathbb{CP}^1$ coset manifold. Plugging in explicit values for the ``color" indices, we obtain the following tree-level amplitude:
\mathcal{A}^{\mathbb{CP}^1}_{\text{tree}}(Z_1,\bar{Z}_2,Z_3,\bar{Z}_4) = f_\pi^{-2}s_{13}\,.
Using this tree-level contact as the seed in our EMU construction, we can directly compute the one- and two-loop amplitudes from \eqn{eq:NLSM1loop} and \eqn{eq:NLSM2loop}. Furthermore, the leading divergences for the loop integrals in $D=2-2\epsilon$ dimensions are the following:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values in for the scattering process $\mathcal{A}(Z_1,\bar{Z}_2,Z_3,\bar{Z}_4) $, we can extract the full leading logarithmic divergence from the one-loop color dressed amplitude in \eqn{eq:NLSM1loop},
\begin{equation}\label{eq:1loopCP1}
\mathcal{A}^{\mathbb{CP}^1}_{\text{1-loop}} = - \left[\frac{if_{\pi}^{-2} }{4\pi \epsilon}\right]f_\pi^{-2}s_{13}+\mathcal{O}(\epsilon^0)\,,
and likewise for the two-loop result upon evaluating the integrals in $D=2-2\epsilon$,

@@

\mathcal{A}^{\mathbb{CP}^1}_{\text{2-loop}} = \frac{1}{2} \left[\frac{if_{\pi}^{-2} }{4\pi \epsilon}\right]^2f_\pi^{-2}s_{13} +\mathcal{O}(\epsilon^{-1})\,.

@@

We write the two-loop amplitude in this suggestive form to emphasize that the logarithmic divergence present at one-loop in \eqn{eq:1loopCP1} appears to exponentiate! More concretely, we have demonstrated through explicit calculation that through two-loop order, the leading divergence of the full color dressed amplitude goes as follows:
\mathcal{A}_{2-2\epsilon}^{\mathbb{CP}^1}\bigg|_{\text{div.}} = \mathcal{A}^{\text{tree}}\left( 1 + \frac{\mathcal{A}^{\text{1-loop}}}{ \mathcal{A}^{\text{tree}}} + \frac{1}{2}\left[\frac{\mathcal{A}^{\text{1-loop}}}{ \mathcal{A}^{\text{tree}}} \right]^2 + \cdots \right)

@@

This iterative form suggests that the IR divergence of the $\mathbb{CP}^1$ model can be neatly extracted from the perturbative series as an overall factor, $e^\Omega$, where $\Omega$ is the divergent piece of the one-loop amplitude. Similar exponentiation of the IR has a long history in the context of gravity amplitudes \cite{Weinberg:1965nx}, for which IR divergences appear due to soft graviton emissions that mediate long range interactions. Considering the established relationship \cite{deWit:1983xhu,deWit:1992cr,deWit:1992wf, Chiodaroli:2014xia} between coset manifolds (target spaces) and supergravity theories, there's a possibility that the exponentiation of the $\mathbb{CP}^1$ model in \eqn{eq:expNLSM} is related to the analogous behavior of supergravity amplitudes \cite{Naculich:2008ew,White:2011yy,DiVecchia:2019myk,DiVecchia:2019kta,Heissenberg:2021tzo}.

In addition to the exponential structure of \eqn{eq:expNLSM} that we have computed for the $\mathbb{CP}^1$ model, there is potentially a parallel story for the planar limit of the chiral NLSM amplitudes. Indeed, if one takes $D\rightarrow 2$ in \eqn{eq:NLSM2bub}, we find that the leading divergence of the double bubble in $D=2-2\epsilon$ is precisely one half the square of the one-loop divergence in \eqn{eq:1loopPionD2}. However, while the cross-box term appearing in \eqn{eq:NLSMostrich} is subleading when $N_c\rightarrow 0$, there are still relevant contributions from the double-boxes appearing in the ostrich integral.

In principle one could add additional operators \cite{Elvang:2018dco} or particle states \cite{Cachazo:2016njl} to \eqn{eq:NLSMLag} that cancel the additional contributions from the ostrich integral, while preserving the soft behavior needed for our EMU construction. Doing so would extend the exponential behavior to the subleading logarithms in the planar limit, similar to what was found in \cite{Anastasiou:2003kj,Bern:2005iz} for planar $\mathcal{N}=4$ sYM in $D=4-2\epsilon$. The iterated structure of planar $\mathcal{N}=4$ has been linked to the integrability of the theory \cite{Sterman:2002qn,Minahan:2002ve,Bena:2003wd,Beisert:2003jj,Beisert:2003jb,Beisert:2003yb,Dolan:2003uh,Arutyunov:2003rg,Ryzhov:2004nz,Frolov:2005iq}. We leave identifying such an extended theory consistent with exponentiation as a compelling direction of future study.

@@

\mathcal{L}^{\text{BI}+\text{CT}} = \mathcal{L}^{\text{BI}} + \frac{\alpha'^4}{(4\pi)^2}\frac{1}{30}\left(\mathcal{O}^{4+}+ a_{({\text{ev.}})}\mathcal{O}^{\text{ev.}}\right)

@@

\mathcal{M}^{\text{BI+CT}}_{\text{1-loop}}\big|_{\alpha'^6} = \frac{1}{2}+ \frac{\alpha'^4}{(4\pi)^2}\frac{1}{30}\left[\scaleIntAvectorODD{}{}{}{}{$\mathcal{T}^{4+}$}{$t_8F^4$}+a_{\text{ev.}}\scaleIntAvectorODD{}{}{}{}{$\mathcal{T}^{\text{ev.}}$}{$t_8F^4$}\right]+\text{perms}

@@

Given the particular numerical value computed in the previous section, we find the evanescent Wilson coefficient must take the following $D$-dependent value:

@@

\mathcal{L}^{\text{BI}+\text{CT}} = \mathcal{L}^{\text{BI}} + \frac{\alpha'^4}{(4\pi)^2}\frac{1}{30}\left[\mathcal{O}^{4+}- \frac{8}{(D-4)}\mathcal{O}^{\text{ev.}}\right]+\mathcal{O}(\alpha'^6)
